### 2016-09-13 sec5_spm.tex
More revisions...

Our customized Apriori Enumerator is presented in Algorithm~\ref{algo:apriori_mining}. Initially, the edges (pairs of objects) in the star constitute the bottom level (Lines~\ref{code:init-start}-\ref{code:init-end}) and invalid candidates are excluded (Line~\ref{code:simp1}). An indicator \emph{level} is used to control the result size for candidate joins. During each iteration (Lines~\ref{code:level-start}-\ref{code:level-ends}), only candidates with object size equals to \emph{level} are generated (Line~\ref{code:join-start}). When two candidates $c_1$ and $c_2$ are joined, the new candidate becomes $c'=\langle c_1.O \cup c_2.O, c_1.T\cap c_2.T\rangle$ (Lines~\ref{code:join}). To check the validity of the candidate, we calculate $\mathtt{sim}(c'.T)$. If its simplified sequence is empty, $c'$ is excluded from the next level (Line~\ref{code:simp2}). This ensures that all the candidates with $P.O\supseteq c'.O$ are pruned. If a candidate cannot generate any new candidate, then
it is directly reported (Lines~\ref{code:output1-start}-\ref{code:output1-end}).
To further improve the performance, we adopt the idea of \textit{forward closure}~\cite{wang2003closet+,pei2000closet} and  aggressively check if the union of all the current candidates form a valid pattern (Lines~\ref{code:fc-checking-start}-\ref{code:fc-checking-end}). If yes, we can  terminate the algorithm early and output the results.

@@

\subsection{Put Everything Together}
We summarize the workflow of SPARE in Figure~\ref{fig:star_partition} as follows. After the parallel clustering in each snapshot, for ease of presentation, we use an aggregated graph $G_A$ to capture the clustering relationship. However, in the implementation of the map phase, there is no need to create $G_A$ in advance. Instead, we simply need to emit the edges within a star to the same reducer. 
%Before sending stars to reducers, we use a simple best-fit strategy to calculate the star allocation. Under the best-fit strategy, the most costly unallocated star is assigned to the most lightly loaded reducers, where we use the edges in the star as a cost estimation.
Each reducer is an Apriori Enumerator. When receiving a star $Sr_i$, the reducer creates initial candidate patterns. Specifically, for each $o \in Sr_i$, a candidate pattern $\langle o,i: e(o,i) \rangle$ is created. Then it enumerates all the valid patterns from the candidate patterns. The pseudocode of SPARE is presented in Algorithm~\ref{algo:spm_overview}.  
In Apache Spark, further optimization can be made by balancing the loads among reducers at runtime. 
Leveraging
Spark's DAG execution engine~\cite{zaharia2012resilient}, a planning phase can be inserted in between map and reduce phases.
With knowing all map results (i.e., star sizes), a simple best-fit planning strategy can be adopted by assigning the most costly unallocated star to the most lightly loaded reducer, where the edges in a star can be used as a cost estimation.

@@

Compared with TRPM, the SPARE framework does not rely on snapshot replication to guarantee correctness. In addition, we can show that the patterns derived from a star partition are unique and there would not be duplicate patterns mined from different star partitions.
\begin{theorem}[Pattern Uniqueness]
\label{LEM:SPM_CORRECT}
Let $Sr_i$ and $Sr_j$ ($i\neq j$) be two star partitions. Let $P_i$ (resp. $P_j$) be 
the patterns discovered from $Sr_i$ (resp. $Sr_j$). 
Then, $\forall p_i \in P_i, \forall p_j \in P_j$, we have $p_i.O \neq p_j.O$.
\end{theorem}
\begin{proof} 
\vspace{-0.5em}
We prove by contradiction. Suppose there exist $p_i \in P_i$ and $p_j \in P_j$ with the same object set. Note that the center vertex of the star is associated with the minimum id. Let $o_i$ and $o_j$ be the center vertexes of the two partitions and we have $o_i=o_j$. However, $P_i$ and $P_j$ are different stars, meaning their center vertexes are different (i.e., $o_i\neq o_j$), leading to a contradiction. 
\vspace{-0.5em}
\end{proof}

>>>>

Our customized Apriori Enumerator is presented in Algorithm~\ref{algo:apriori_mining}. Initially, the edges (pairs of objects) in the star constitute the bottom level (Lines~\ref{code:init-start}-\ref{code:init-end}) and invalid candidates are excluded (Line~\ref{code:simp1}). An indicator \emph{level} is used to control the result size for candidate joins. During each iteration (Lines~\ref{code:level-start}-\ref{code:level-ends}), only candidates with object size equals to \emph{level} are generated (Line~\ref{code:join-start}). When two candidates $c_1$ and $c_2$ are joined, the new candidate becomes $c'=\langle c_1.O \cup c_2.O, c_1.T\cap c_2.T\rangle$ (Line~\ref{code:join}). To check the validity of the candidate, we calculate $\mathtt{sim}(c'.T)$. If its simplified sequence is empty, $c'$ is excluded from the next level (Line~\ref{code:simp2}). This ensures that all the candidates with $P.O\supseteq c'.O$ are pruned. If a candidate cannot generate any new candidate, then
it is directly reported (Lines~\ref{code:output1-start}-\ref{code:output1-end}).
To further improve the performance, we adopt the idea of \textit{forward closure}~\cite{wang2003closet+,pei2000closet} and  aggressively check if the union of all the current candidates form a valid pattern (Lines~\ref{code:fc-checking-start}-\ref{code:fc-checking-end}). If yes, we can  terminate the algorithm early and output the results.

@@

\subsection{Put Everything Together}
We summarize the workflow of SPARE in Figure~\ref{fig:star_partition} as follows. After the parallel clustering in each snapshot, for ease of presentation, we use an aggregated graph $G_A$ to capture the clustering relationship. However, in the implementation of the map phase, there is no need to create $G_A$ in advance. Instead, we simply need to emit the edges within a star to the same reducer. 
%Before sending stars to reducers, we use a simple best-fit strategy to calculate the star allocation. Under the best-fit strategy, the most costly unallocated star is assigned to the most lightly loaded reducers, where we use the edges in the star as a cost estimation.
Each reducer is an Apriori Enumerator. When receiving a star $Sr_i$, the reducer creates initial candidate patterns. Specifically, for each $o \in Sr_i$, a candidate pattern $\langle o,i: e(o,i) \rangle$ is created. Then it enumerates all the valid patterns from the candidate patterns. The pseudocode of SPARE is presented in Algorithm~\ref{algo:spm_overview}. 
In our implementation of SPARE on Spark~\cite{zaharia2012resilient}, we take advantage of Spark features to achieve better workload balance. In particular,
we utilize Spark DAG execution engine to inject a planning phase between map and reduce phases. By knowing all map results (i.e., star sizes), a simple best-fit strategy is adopted which assigns the most costly unallocated star to the most lightly loaded reducer, where the edges in a star are used as cost estimations. We also leverage Spark in-memory cache to avoid recomputing all stars after the planning phase.
% 
%In Apache Spark, further optimization can be made by balancing the loads among reducers at runtime. 
%Leveraging
%Spark's DAG execution engine~\cite{zaharia2012resilient}, a planning phase can be inserted in between map and reduce phases.
%With knowing all map results (i.e., star sizes), a simple best-fit planning strategy can be adopted by assigning the most costly unallocated star to the most lightly loaded reducer, where the edges in a star can be used as a cost estimation.

@@

Compared with TRPM, the SPARE framework does not rely on snapshot replication to guarantee correctness. In addition, we can show that the patterns derived from a star partition are unique and there would not be duplicate patterns mined from different star partitions.
\begin{theorem}[Pattern Uniqueness]
\label{LEM:SPM_CORRECT}
Let $Sr_i$ and $Sr_j$ ($i\neq j$) be two star partitions. Let $P_i$ (resp. $P_j$) be 
the patterns discovered from $Sr_i$ (resp. $Sr_j$). 
Then, $\forall p_i \in P_i, \forall p_j \in P_j$, we have $p_i.O \neq p_j.O$.
\end{theorem}
\begin{proof} 
\vspace{-0.5em}
We prove by contradiction. Suppose there exist $p_i \in P_i$ and $p_j \in P_j$ with the same object set. Note that the center vertex of the star is associated with the minimum id. Let $o_i$ and $o_j$ be the center vertexes of the two partitions and we have $o_i=o_j$. However, $P_i$ and $P_j$ are from different stars, meaning their center vertexes are different (i.e., $o_i\neq o_j$), leading to a contradiction. 
\vspace{-0.5em}
\end{proof}

### 2016-09-13 sec6_experiment.tex
More revisions...

\textbf{Environment Setup}: We use Yarn\footnote{\url{http://hadoop.apache.org/docs/current/hadoop-yarn/hadoop-yarn-site/YARN.html}} to manage our cluster. We pick one machine as Yarn's master node, and for each of the remaining machines, we reserve one core and $2$GB memory for Yarn processes. We deploy our GCMP detector on Apache Spark 1.5.2~\cite{zaharia2012resilient} with the remaining $11$ nodes as the computing nodes.
To fully utilize the computing resources, we configure each node to run five executors, each taking three cores and $5$GB memory. In Spark, one of the $55$ executors is taken as the Application Master for coordination, therefore our setting results in $54$ executors.  We set the number of partitions to be $486$ to fully utilize the multi-threading feature of every core.
All our implementations as well as cluster setups are publicly available% in Github
\footnote{\url{https://github.com/fanqi1909/TrajectoryMining/}}.

\textbf{Datasets}: We use three real trajectory datasets in different application scenarios:
\begin{itemize}[nosep]
\item{Shopping}\footnote{\url{http://www.irc.atr.jp/crest2010_HRI/ATC_dataset/}}: The dataset contains
  trajectories of visitors in the ATC shopping center in Osaka. To better capture the indoor activities, the visitor locations are sampled every half second, resulting in $13,183$ long trajectories. 
\item{GeoLife}~\footnote{\url{http://research.microsoft.com/en-us/projects/geolife/}}: The dataset essentially keeps all the travel records of 182 users for a period
of over three years, including multiple kinds of transportation modes (walking, driving and taking public
transportation). For each user, the GPS information is collected periodically and 91 percent of the trajectories
are sampled every 1 to 5 seconds.
\item{Taxi}~\footnote{Taxi is our proprietary dataset}: The dataset tracks the trajectories of $15,054$ taxies in Singapore. For each taxi, the GPS information are continually collected for one entire month with the sampling rate around 30 seconds.
\end{itemize}

@@

%\vspace{-5mm}
\begin{table}[h]
\small
\begin{tabular}{c|l|p{3cm}}
\hline 
\textbf{Parameter} & \textbf{Meaning} & \textbf{Values} \\ 
\hline 
M & min  objects &  5, 10,  \textbf{15}, 20, 25 \\ 
\hline 
K & min duration & 120, 150, \textbf{180}, 210, 240 \\ 
\hline 
L & min local duration & 10, 20, \textbf{30}, 40,50 \\ 
\hline 
G & max gap & 10, 15, \textbf{20}, 25, 30 \\ 
\hline
$O_r$ & ratio of objects & 20\%,40\%, 60\%, 80\%, \textbf{100\%} \\ \hline
$T_r$ & ratio of snapshots & 20\%,40\%, 60\%, 80\%, \textbf{100\%} \\ \hline
N & number of machines & 1, 3, 5, 7, 9, \textbf{11}\\ 
\hline 
\end{tabular} 
 \vspace{-0.5em}
\caption{Parameters and their default values.}
 \vspace{-0.5em}
\label{tbl:parameters}
\end{table}

@@

\textbf{Varying $K$}: The performance with increasing $K$ is shown in Figure~\ref{exp:performance_vary} (b),(h),(n).  SPARE tends to run faster, whereas the performance of TRPM degrades dramatically. This is caused by the \emph{sequence simplification} procedure in SPARE, which can prune many candidates with large $K$. However, the line sweep algorithm in TRPM does not utilize such property for pruning. It takes longer time because more replicated data has to be handled in each partition.

@@

\textbf{Varying $O_r$}: Figures~\ref{exp:performance_vary} (e),(k),(q) present the performances with increasing number of moving objects. Both TRPM and SPARE take longer time to find patterns in a larger database. We can see that the performance gap between SPARE and TRPM is widened as more objects are involved, which shows SPARE is more scalable.

@@

\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{\textbf{Dataset}} & \textbf{Method} % & \textbf{Time} 
& \textbf{Vcore-seconds} & \textbf{Memory} \\ \hline
\multirow{2}{*}{Shopping}              & TRPM           % & 597
                               & 90,859                 & 10,019               \\ %\cline{2-5} 
                                       & SPARE           %& 237
                                                                      & 33,638                 & 8,613                \\ \hline
\multirow{2}{*}{Geolife}               & TRPM            %& 747 
                              & 106,428                & 18,454               \\ %\cline{2-5} 
                                       & SPARE           %& 341 
                                                                     & 35,343                 & 14,369               \\ \hline
\multirow{2}{*}{Taxi}                  & TRPM            %& 3,648  
                           & 503,460                & 51,691               \\ %\cline{2-5} 
                                       & SPARE           %& 580  
                                                                    & 68,580                & 35,912               \\ \hline
\end{tabular}
 \vspace{-0.5em}
\caption{Resources taken for TRPM and SPARE. Vcore-seconds is the aggregate of time spent in each core. Memory is the actual size of RDDs.}
\vspace{-0.5em}
\label{tbl:resource}
\end{table}

@@

\subsubsection{Power of sequence simplification}
To study the power of sequence simplification,
we collect two types of statistics: (1) the number of pairs that
are shuffled to the reducers and (2) the number of pairs that
are fed to the Apirori Enumerator. The difference between these two values is the number of size-$2$ candidates pruned by the sequence simplification.
The results in Table~\ref{tbl:pruning} show that the \emph{sequence simplification} is very powerful and eliminates nearly 90 percent of the object pairs, which significantly reduces the overhead of subsequent Apriori enumerator.

@@

\subsubsection{Load balance}
To study the effect of load balance in the SPARE framework, we use random task allocation (the default setting of Spark) as a baseline, denoted by SPARE-RD, and compare it with our best-fit method. In best-fit, the largest unassigned star is allocated to the currently most lightly loaded reducer.
Figure~\ref{exp:wl} shows the breakdown of the costs in the map-reduce stages 
for SPARE and SPARE-RD. %\revised{with TRPM as comparison}. 
We observe that the map and shuffle time of SPARE and SPARE-RD are identical. The difference is that SPARE incurs an additional overhead to generate an allocation plan for load balance (around $4\%$ of the total cost), resulting in significant savings in the reduce stage (around $20\%$ of the total cost). 
Meanwhile, both SPARE and SPARE-RD outperform TRPM in each phase. This indicates the 
efficiency of the star partition and apriori enumeration.
%\revised{We also report the cost summaries (i.e., longest, shortest and standard deviation) of all jobs for the three methods in Table~\ref{tbl:strags}. As the table shows, SPARE produces more balanced jobs
%than SPARE-RD which verify the effectiveness of the best-fit strategy. In addition, TRPM 
%maintains a low deviation because it uses equal size (i.e., $\eta$) partitioning. 
%However, TRPM still takes much longer time than SPARE algorithms.}
We also report the cost  for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.
We also report the cost of the longest job (Max) and the standard deviation (Std. Dev.)
%\emph{straggler}, i.e., the longest job, and the standard deviation (Std. Dev.) 
for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.

>>>>

\textbf{Environment Setup}: We use Yarn\footnote{\url{http://hadoop.apache.org/docs/current/hadoop-yarn/hadoop-yarn-site/YARN.html}} to manage our cluster. We pick one machine as Yarn's master node, and for each of the remaining machines, we reserve one core and $2$GB memory for Yarn processes. We deploy our GCMP detector on Apache Spark 1.5.2 with the remaining $11$ nodes as the computing nodes.
To fully utilize the computing resources, we configure each node to run five executors, each taking three cores and $5$GB memory. In Spark, one of the $55$ executors is taken as the Application Master for coordination, therefore our setting results in $54$ executors.  We set the number of partitions to be $486$ to fully utilize the multi-threading feature of every core.
All our implementations as well as cluster setups are publicly available% in Github
\footnote{\url{https://github.com/fanqi1909/TrajectoryMining/}, we have also experimented with a query-based TRPM using Spark-SQL 2.0.0 window function. 
We find that Spark-SQL fails to execute the query-based TRPM in parallel, which results in a 120x performance slowdown compared to mapreduce-based TRPM.
%
%it runs over 120 times slower than the mapreduce-based TRPM. 
Thus we only report the performance of mapreduce-based TRPM in this paper.}.

\textbf{Datasets}: We use three real trajectory datasets in different application scenarios:
\begin{itemize}[nosep]
\item{Shopping}\footnote{\url{http://www.irc.atr.jp/crest2010_HRI/ATC_dataset/}}: The dataset contains
  trajectories of visitors in the ATC shopping center in Osaka. To better capture the indoor activities, the visitor locations are sampled every half second, resulting in $13,183$ long trajectories. 
\item{GeoLife}\footnote{\url{http://research.microsoft.com/en-us/projects/geolife/}}: The dataset essentially keeps all the travel records of 182 users for a period
of over three years, including multiple kinds of transportation modes (walking, driving and taking public
transportation). For each user, the GPS information is collected periodically and 91 percent of the trajectories
are sampled every 1 to 5 seconds.
\item{Taxi}\footnote{Taxi is our proprietary dataset}: The dataset tracks the trajectories of $15,054$ taxies in Singapore. For each taxi, the GPS information are continually collected for one entire month with the sampling rate around 30 seconds.
\end{itemize}

@@

\begin{table}[h]
\small
\begin{tabular}{c|l|p{3cm}}
\hline 
\textbf{Parameter} & \textbf{Meaning} & \textbf{Values} \\ 
\hline 
M & min  objects &  5, 10,  \textbf{15}, 20, 25 \\ 
\hline 
K & min duration & 120, 150, \textbf{180}, 210, 240 \\ 
\hline 
L & min local duration & 10, 20, \textbf{30}, 40,50 \\ 
\hline 
G & max gap & 10, 15, \textbf{20}, 25, 30 \\ 
\hline
$O_r$ & ratio of objects & 20\%,40\%, 60\%, 80\%, \textbf{100\%} \\ \hline
$T_r$ & ratio of snapshots & 20\%,40\%, 60\%, 80\%, \textbf{100\%} \\ \hline
N & number of machines & 1, 3, 5, 7, 9, \textbf{11}\\ 
\hline 
\end{tabular} 
 \vspace{-0.5em}
\caption{Parameters and their default values.}
 \vspace{-0.5em}
\label{tbl:parameters}
\end{table}

@@

\textbf{Varying $K$}: The performance with increasing $K$ is shown in Figures~\ref{exp:performance_vary} (b),(h),(n).  SPARE tends to run faster, whereas the performance of TRPM degrades dramatically. This is caused by the \emph{sequence simplification} procedure in SPARE, which can prune many candidates with large $K$. However, the line sweep algorithm in TRPM does not utilize such property for pruning. It takes longer time because more replicated data has to be handled in each partition.

@@

\textbf{Varying $O_r$}:Figures~\ref{exp:performance_vary} (e),(k),(q) present the performances with increasing number of moving objects. Both TRPM and SPARE take longer time to find patterns in a larger database. We can see that the performance gap between SPARE and TRPM is widened as more objects are involved, which shows SPARE is more scalable.

@@

\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
\multicolumn{1}{|c|}{\textbf{Dataset}} & \textbf{Method} % & \textbf{Time} 
& \textbf{Vcore-seconds} & \textbf{Memory} \\ \hline
\multirow{2}{*}{Shopping}              & TRPM           % & 597
                               & 90,859                 & 10,019               \\ %\cline{2-5} 
                                       & SPARE           %& 237
                                                                      & 33,638                 & 8,613                \\ \hline
\multirow{2}{*}{Geolife}               & TRPM            %& 747 
                              & 106,428                & 18,454               \\ %\cline{2-5} 
                                       & SPARE           %& 341 
                                                                     & 35,343                 & 14,369               \\ \hline
\multirow{2}{*}{Taxi}                  & TRPM            %& 3,648  
                           & 503,460                & 51,691               \\ %\cline{2-5} 
                                       & SPARE           %& 580  
                                                                    & 68,580                & 35,912               \\ \hline
\end{tabular}
 \vspace{-0.5em}
\caption{Resources taken for TRPM and SPARE. Vcore-seconds is the aggregate of time spent in each core. Memory is the actual size (in MB) of RDDs.}
\vspace{-0.5em}
\label{tbl:resource}
\end{table}

@@

\subsubsection{Power of sequence simplification}
To study the power of \emph{Sequence Simplification} (SS),
we collect two types of statistics: (1) the number of pairs that
are shuffled to the reducers and (2) the number of pairs that
are fed to the Apirori Enumerator. 
Their difference %is
%The difference between these two values 
is the number of size-$2$ candidates pruned by SS.
The results in Table~\ref{tbl:pruning} show that %the \emph{sequence simplification} 
SS is very powerful and eliminates nearly 90 percent of the object pairs, 
which significantly reduces the overhead of %subsequent
the Apriori enumerator.
%Notice that SS is
%essential for the monotonicity pruning of Apriori enumerator. 
In fact, without SS Apriori cannot
finish in five hours.

@@

\subsubsection{Load balance}
To study the effect of load balance in the SPARE framework, we use random task allocation (the default setting of Spark) as a baseline, denoted by SPARE-RD, and compare it with our best-fit method. In best-fit, the largest unassigned star is allocated to the currently most lightly loaded reducer.
Figure~\ref{exp:wl} shows the breakdown of the costs in the mapreduce stages 
for SPARE and SPARE-RD. %\revised{with TRPM as comparison}. 
We observe that the map and shuffle time of SPARE and SPARE-RD are identical. The difference is that SPARE incurs an additional overhead to generate an allocation plan for load balance (around $4\%$ of the total cost), resulting in significant savings in the reduce stage (around $20\%$ of the total cost). 
Meanwhile, both SPARE and SPARE-RD outperform TRPM in each phase. This indicates the 
efficiency of the star partition and apriori enumeration.
%\revised{We also report the cost summaries (i.e., longest, shortest and standard deviation) of all jobs for the three methods in Table~\ref{tbl:strags}. As the table shows, SPARE produces more balanced jobs
%than SPARE-RD which verify the effectiveness of the best-fit strategy. In addition, TRPM 
%maintains a low deviation because it uses equal size (i.e., $\eta$) partitioning. 
%However, TRPM still takes much longer time than SPARE algorithms.}
%We also report the cost  for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.
We also report the cost of the longest job (Max) and the standard deviation (Std. Dev.)
%\emph{straggler}, i.e., the longest job, and the standard deviation (Std. Dev.) 
for all jobs in Table~\ref{tbl:strags}, whose results clearly verify the effectiveness of our allocation strategy for load balance.
